---
title: "R Basics for Bioinformatics Scientists"
subtitle: "From Variables to Statistical Analysis with the Tidyverse"
author: "Portfolio 2026"
date: today
format:
  html:
    theme: flatly
    toc: true
    toc-depth: 3
    code-fold: show
    code-tools: true
    self-contained: true
    fig-align: center
execute:
  warning: false
  message: false
---

# Part 1: R Fundamentals

## 1.1 Variables and Data Types

R uses `<-` as its primary assignment operator. The `=` sign also works but `<-` is
the idiomatic choice.

```{r}
#| label: variables-basics

# Assignment
gene_name <- "BRCA1"
expression_value <- 42.7
is_upregulated <- TRUE
sample_count <- 6L          # L suffix forces integer storage

# Check types
class(gene_name)
class(expression_value)
class(is_upregulated)
class(sample_count)
typeof(sample_count)
```

### Special Values

R has several special values you will encounter regularly when working with
biological data:

```{r}
#| label: special-values

NA                           # Missing value (unknown)
NULL                         # Empty / absent object
NaN                          # Not a number (0 / 0)
Inf                          # Infinity (1 / 0)

# Testing for special values
is.na(NA)
is.null(NULL)
is.nan(NaN)
is.finite(Inf)
```

### Type Coercion

R converts types automatically following the hierarchy:
logical → integer → numeric → character.

```{r}
#| label: coercion

# Automatic coercion
c(TRUE, 1L, 3.14, "hello")  # everything becomes character

# Explicit coercion
as.numeric("42.5")
as.integer(3.7)              # truncates, does NOT round
as.character(100)
as.logical(0)                # 0 → FALSE, non-zero → TRUE
```

## 1.2 Vectors

Vectors are the most fundamental data structure in R. Every element must be the
same type.

```{r}
#| label: vectors

# Creating vectors
gene_lengths <- c(1863, 5592, 2487, 945, 3210)
samples <- c("ctrl_1", "ctrl_2", "treat_1", "treat_2")
log_fc <- seq(-2, 2, by = 0.5)
replicates <- rep(c("A", "B", "C"), times = 3)

# Vectorised arithmetic
gene_lengths_kb <- gene_lengths / 1000
gene_lengths_kb

# Logical comparisons (returns logical vector)
long_genes <- gene_lengths > 2000
long_genes

# Indexing
gene_lengths[1]              # first element
gene_lengths[c(1, 3)]        # multiple elements
gene_lengths[-2]             # exclude second element
gene_lengths[long_genes]     # logical indexing
```

### Named Vectors

```{r}
#| label: named-vectors

expression <- c(BRCA1 = 12.3, TP53 = 8.7, MYC = 45.1, EGFR = 22.0)
expression["MYC"]

# Useful functions
length(expression)
names(expression)
sort(expression)
which.max(expression)
sum(expression > 15)         # count elements meeting a condition
```

## 1.3 Lists

Lists can hold elements of different types and lengths — useful for storing
mixed results (e.g., from a statistical test).

```{r}
#| label: lists

experiment <- list(
  name       = "RNA-seq time course",
  organism   = "Homo sapiens",
  timepoints = c(0, 6, 12, 24),
  n_samples  = 12,
  paired     = TRUE
)

# Access elements
experiment$organism          # by name with $
experiment[["timepoints"]]   # by name with [[]]
experiment[[4]]              # by position

# str() shows the structure
str(experiment)
```

## 1.4 Matrices

Matrices are 2D structures where every element shares the same type. They
frequently represent gene expression tables before conversion to data frames.

```{r}
#| label: matrices

# Create a small expression matrix (genes × samples)
expr_mat <- matrix(
  c(5.2, 3.1, 8.4, 6.7,
    2.0, 1.5, 7.3, 5.9,
    9.1, 8.8, 4.2, 3.5),
  nrow = 3, byrow = TRUE
)
rownames(expr_mat) <- c("GeneA", "GeneB", "GeneC")
colnames(expr_mat) <- c("S1", "S2", "S3", "S4")
expr_mat

# Indexing
expr_mat["GeneA", ]          # one row
expr_mat[, "S2"]             # one column
expr_mat[1:2, 3:4]           # sub-matrix

# Row/column operations
rowMeans(expr_mat)
apply(expr_mat, 1, sd)       # row-wise standard deviation
t(expr_mat)                  # transpose
```

## 1.5 Data Frames

Data frames are the workhorse for tabular data in R. Each column can hold a
different type.

```{r}
#| label: data-frames

# Create from scratch
samples_df <- data.frame(
  sample_id  = c("S1", "S2", "S3", "S4", "S5", "S6"),
  group      = c("ctrl", "ctrl", "ctrl", "treat", "treat", "treat"),
  rin_score  = c(8.2, 7.9, 8.5, 7.1, 8.0, 8.3),
  lib_size   = c(25e6, 30e6, 28e6, 32e6, 27e6, 29e6)
)

# Inspection functions
str(samples_df)
summary(samples_df)
head(samples_df, 3)
dim(samples_df)
nrow(samples_df)
ncol(samples_df)
```

### Built-in Datasets

R ships with many practice datasets:

```{r}
#| label: builtin-datasets

head(mtcars, 5)
head(iris, 5)
```

### Tibbles (Preview)

The tidyverse uses **tibbles** — a modern take on data frames with better
printing and stricter subsetting. We will use them extensively in Part 2.

```{r}
#| label: tibble-preview

library(tibble)
as_tibble(mtcars)
```

## 1.6 Control Flow

### Conditionals

```{r}
#| label: conditionals

pvalue <- 0.03

if (pvalue < 0.001) {
  significance <- "***"
} else if (pvalue < 0.01) {
  significance <- "**"
} else if (pvalue < 0.05) {
  significance <- "*"
} else {
  significance <- "ns"
}
significance

# Vectorised conditional
log_fc_vals <- c(-2.1, 0.3, 1.8, -0.1, 3.5)
direction <- ifelse(log_fc_vals > 0, "up", "down")
direction
```

### Loops

```{r}
#| label: loops

# For loop — iterate over gene names
genes <- c("BRCA1", "TP53", "MYC")
for (gene in genes) {
  cat("Processing:", gene, "\n")
}

# Collecting results in a pre-allocated vector
gc_content <- c(0.42, 0.55, 0.61)
names(gc_content) <- genes

high_gc <- character(0)
for (i in seq_along(gc_content)) {
  if (gc_content[i] > 0.5) {
    high_gc <- c(high_gc, names(gc_content)[i])
  }
}
high_gc
```

> **Note**: In idiomatic R — especially with the tidyverse — vectorised
> operations and `dplyr` verbs replace most loops.

## 1.7 Functions

```{r}
#| label: functions

# Calculate GC content of a DNA sequence
gc_content <- function(sequence) {
  bases <- strsplit(toupper(sequence), "")[[1]]
  gc <- sum(bases %in% c("G", "C"))
  gc / length(bases)
}

gc_content("ATGCGCTA")
gc_content("AAATTTAAA")
```

### Default Arguments and Lambda Syntax

```{r}
#| label: functions-advanced

# Default arguments
normalize <- function(x, method = "zscore") {
  if (method == "zscore") {
    return((x - mean(x)) / sd(x))
  }
  if (method == "minmax") {
    return((x - min(x)) / (max(x) - min(x)))
  }
  stop("Unknown method: ", method)
}

vals <- c(10, 20, 30, 40, 50)
normalize(vals)
normalize(vals, method = "minmax")

# Lambda (anonymous) function — R 4.1+ syntax
sapply(c("ATG", "GCGC", "AATT"), \(s) nchar(s))
```


# Part 2: Loading Data with the Tidyverse

## 2.1 Setup

```{r}
#| label: tidyverse-setup

library(tidyverse)    # loads dplyr, ggplot2, tidyr, readr, purrr, stringr, forcats, tibble
library(readxl)
library(writexl)
```

The `tidyverse` is a collection of packages that share a common design
philosophy: tidy data, consistent function signatures, and the pipe operator.

### The Pipe Operator

The pipe passes the result of the left side as the **first argument** of the
right side, making chains of transformations readable:
```r
# Without pipe
summary(filter(select(data, col1, col2), col1 > 5))

# With pipe — read top-to-bottom
data |>
  select(col1, col2) |>
  filter(col1 > 5) |>
  summary()
```

R 4.1+ provides the native pipe `|>`. The tidyverse also offers `%>%` from
`magrittr`; both work identically for the vast majority of use cases.

## 2.2 Creating Sample Data Files

We create small example files inside a temporary directory so the tutorial is
fully self-contained.

```{r}
#| label: create-sample-files

# Simulated gene expression experiment
set.seed(42)
experiment_data <- tibble(
  sample_id   = paste0("S", 1:20),
  group       = rep(c("control", "treatment"), each = 10),
  gene_BRCA1  = c(rnorm(10, mean = 5, sd = 1), rnorm(10, mean = 8, sd = 1.2)),
  gene_TP53   = c(rnorm(10, mean = 12, sd = 2), rnorm(10, mean = 10, sd = 1.8)),
  gene_MYC    = c(rnorm(10, mean = 3, sd = 0.5), rnorm(10, mean = 6, sd = 1)),
  batch       = rep(c("batch_1", "batch_2"), times = 10),
  rin_score   = round(runif(20, min = 6.5, max = 9.5), 1)
)

# Sample metadata
sample_meta <- tibble(
  sample_id = paste0("S", 1:20),
  age       = sample(25:65, 20, replace = TRUE),
  sex       = sample(c("M", "F"), 20, replace = TRUE),
  site      = sample(c("site_A", "site_B", "site_C"), 20, replace = TRUE)
)

tmp <- tempdir()

# Write in different formats
write_csv(experiment_data, file.path(tmp, "expression.csv"))
write_tsv(experiment_data, file.path(tmp, "expression.tsv"))
write_xlsx(experiment_data, file.path(tmp, "expression.xlsx"))
write_csv(sample_meta, file.path(tmp, "metadata.csv"))
```

## 2.3 Reading Data

### CSV with `readr::read_csv()`

```{r}
#| label: read-csv

expr <- read_csv(file.path(tmp, "expression.csv"))
expr
```

Specify column types explicitly to avoid surprises:

```{r}
#| label: read-csv-coltypes

expr <- read_csv(
  file.path(tmp, "expression.csv"),
  col_types = cols(
    sample_id  = col_character(),
    group      = col_factor(levels = c("control", "treatment")),
    gene_BRCA1 = col_double(),
    gene_TP53  = col_double(),
    gene_MYC   = col_double(),
    batch      = col_character(),
    rin_score  = col_double()
  )
)
```

### TSV with `readr::read_tsv()`

```{r}
#| label: read-tsv

expr_tsv <- read_tsv(file.path(tmp, "expression.tsv"))
expr_tsv
```

### Excel with `readxl::read_excel()`

```{r}
#| label: read-excel

expr_xl <- read_excel(file.path(tmp, "expression.xlsx"))
expr_xl

# List available sheets
excel_sheets(file.path(tmp, "expression.xlsx"))
```

### Comparison with Base R

| Feature | `read_csv()` | `read.csv()` |
|---------|-------------|--------------|
| Returns | tibble | data.frame |
| Speed | faster | slower |
| Column names | keeps as-is | converts `.` |
| Strings | character | factor (old R) |
| Progress bar | yes (large files) | no |


# Part 3: Data Wrangling with dplyr and tidyr

```{r}
#| label: load-penguins

library(palmerpenguins)
penguins_clean <- penguins |> drop_na()
penguins_clean
```

## 3.1 Core dplyr Verbs

### `filter()` — Select Rows

```{r}
#| label: filter

# Single condition
penguins_clean |> filter(species == "Adelie")

# Multiple conditions
penguins_clean |>
  filter(species == "Gentoo", body_mass_g > 5500)

# Using %in%
penguins_clean |>
  filter(island %in% c("Biscoe", "Dream"))

# Using between()
penguins_clean |>
  filter(between(bill_length_mm, 40, 50))
```

### `select()` — Choose Columns

```{r}
#| label: select

# By name
penguins_clean |> select(species, bill_length_mm, body_mass_g)

# With helpers
penguins_clean |> select(starts_with("bill"))
penguins_clean |> select(where(is.numeric))

# Drop columns
penguins_clean |> select(-year, -island)
```

### `mutate()` — Add or Modify Columns

```{r}
#| label: mutate

penguins_clean |>
  mutate(
    body_mass_kg  = body_mass_g / 1000,
    bill_ratio    = bill_length_mm / bill_depth_mm,
    size_category = case_when(
      body_mass_g > 5000 ~ "large",
      body_mass_g > 3500 ~ "medium",
      TRUE               ~ "small"
    )
  ) |>
  select(species, body_mass_kg, bill_ratio, size_category)
```

### `summarize()` and `group_by()`

```{r}
#| label: summarize

# Ungrouped
penguins_clean |>
  summarize(
    n           = n(),
    mean_mass   = mean(body_mass_g),
    sd_mass     = sd(body_mass_g),
    median_bill = median(bill_length_mm)
  )

# Grouped by species
penguins_clean |>
  group_by(species) |>
  summarize(
    n         = n(),
    mean_mass = mean(body_mass_g),
    sd_mass   = sd(body_mass_g),
    .groups   = "drop"
  )

# Grouped by species and sex
penguins_clean |>
  group_by(species, sex) |>
  summarize(
    mean_flipper = mean(flipper_length_mm),
    .groups = "drop"
  )
```

### `arrange()` — Sort Rows

```{r}
#| label: arrange

penguins_clean |>
  group_by(species) |>
  summarize(mean_mass = mean(body_mass_g), .groups = "drop") |>
  arrange(desc(mean_mass))
```

### `count()` and `distinct()`

```{r}
#| label: count-distinct

penguins_clean |> count(species, island, sort = TRUE)
penguins_clean |> distinct(species, island)
```

### `slice_*()` Functions

```{r}
#| label: slice

# Top 3 heaviest penguins per species
penguins_clean |>
  group_by(species) |>
  slice_max(body_mass_g, n = 3)

# Random sample
penguins_clean |> slice_sample(n = 5)
```

### `across()` — Apply Functions to Multiple Columns

```{r}
#| label: across

penguins_clean |>
  group_by(species) |>
  summarize(
    across(where(is.numeric), list(mean = mean, sd = sd)),
    .groups = "drop"
  )
```

## 3.2 Joins

Joins combine two tables by matching rows on shared columns — essential for
linking expression data to sample annotations.

```{r}
#| label: joins-setup

gene_info <- tibble(
  gene_id    = c("ENSG001", "ENSG002", "ENSG003", "ENSG004"),
  gene_name  = c("BRCA1", "TP53", "MYC", "UNKNOWN"),
  chromosome = c("chr17", "chr17", "chr8", "chr1")
)

expression <- tibble(
  gene_id   = c("ENSG001", "ENSG002", "ENSG003", "ENSG005"),
  sample    = c("S1", "S1", "S1", "S1"),
  log2_expr = c(5.2, 8.1, 3.7, 6.0)
)
```

```{r}
#| label: joins

# Keep all rows from the left table
left_join(expression, gene_info, by = "gene_id")

# Keep only rows present in BOTH tables
inner_join(expression, gene_info, by = "gene_id")

# Keep all rows from both tables
full_join(expression, gene_info, by = "gene_id")

# Rows in expression that have NO match in gene_info
anti_join(expression, gene_info, by = "gene_id")
```

## 3.3 Reshaping with tidyr

### `pivot_longer()` — Wide to Long

Many ggplot2 geoms and statistical functions require **long** (tidy) format
where each observation is a row.

```{r}
#| label: pivot-longer

# Our expression data has genes as columns — make it long
expr_long <- expr |>
  pivot_longer(
    cols      = starts_with("gene_"),
    names_to  = "gene",
    values_to = "expression",
    names_prefix = "gene_"
  )
expr_long
```

### `pivot_wider()` — Long to Wide

```{r}
#| label: pivot-wider

expr_long |>
  pivot_wider(
    names_from  = gene,
    values_from = expression,
    names_prefix = "gene_"
  )
```

### `separate()` and `unite()`

```{r}
#| label: separate-unite

genomic_coords <- tibble(
  locus = c("chr1:1000-2000", "chr5:3400-4500", "chr17:7500-8900")
)

# Split into components
genomic_coords |>
  separate(locus, into = c("chr", "start", "end"), sep = "[:-]")

# Combine columns
tibble(chr = c("chr1", "chr5"), start = c(1000, 3400), end = c(2000, 4500)) |>
  unite("locus", chr, start, end, sep = ":")
```

### Handling Missing Values

```{r}
#| label: missing-values

df_with_na <- tibble(
  gene  = c("A", "B", "C", "D"),
  expr1 = c(5.2, NA, 3.1, 7.0),
  expr2 = c(NA, 4.3, 2.8, 6.5)
)

df_with_na |> drop_na()                          # remove rows with any NA
df_with_na |> drop_na(expr1)                     # only check expr1
df_with_na |> replace_na(list(expr1 = 0, expr2 = 0))
```


# Part 4: Plotting with ggplot2

## 4.1 The Grammar of Graphics

ggplot2 builds plots in layers:

1. **Data** — the tibble/data frame
2. **Aesthetics** (`aes()`) — map variables to visual properties
3. **Geoms** — geometric objects (points, lines, bars, …)
4. **Scales** — control how values map to aesthetics
5. **Facets** — split into sub-plots
6. **Theme** — control non-data appearance

```{r}
#| label: first-plot
#| fig-width: 7
#| fig-height: 5

ggplot(penguins_clean, aes(x = bill_length_mm, y = body_mass_g, color = species)) +
  geom_point(alpha = 0.7, size = 2) +
  labs(
    title = "Penguin Body Mass vs Bill Length",
    x = "Bill Length (mm)",
    y = "Body Mass (g)",
    color = "Species"
  ) +
  theme_minimal()
```

## 4.2 Common Geom Types

### Scatter Plot — `geom_point()`

```{r}
#| label: geom-point
#| fig-width: 7
#| fig-height: 5

ggplot(penguins_clean, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species, shape = sex), size = 2, alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, color = "grey30", linewidth = 0.8) +
  theme_minimal()
```

### Bar Charts — `geom_bar()` and `geom_col()`

```{r}
#| label: geom-bar
#| fig-width: 6
#| fig-height: 4

# geom_bar counts for you
ggplot(penguins_clean, aes(x = species, fill = island)) +
  geom_bar(position = "dodge") +
  labs(title = "Penguin Counts by Species and Island") +
  theme_minimal()
```

```{r}
#| label: geom-col
#| fig-width: 6
#| fig-height: 4

# geom_col for pre-computed values
penguins_clean |>
  group_by(species) |>
  summarize(mean_mass = mean(body_mass_g), .groups = "drop") |>
  ggplot(aes(x = species, y = mean_mass, fill = species)) +
  geom_col(show.legend = FALSE) +
  labs(y = "Mean Body Mass (g)") +
  theme_minimal()
```

### Histogram and Density

```{r}
#| label: histogram-density
#| fig-width: 8
#| fig-height: 4

library(patchwork)

p1 <- ggplot(penguins_clean, aes(x = body_mass_g, fill = species)) +
  geom_histogram(bins = 30, alpha = 0.6, position = "identity") +
  labs(title = "Histogram") +
  theme_minimal()

p2 <- ggplot(penguins_clean, aes(x = body_mass_g, fill = species)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density") +
  theme_minimal()

p1 + p2
```

### Box Plot, Violin Plot, and Jitter

```{r}
#| label: boxplot-violin
#| fig-width: 9
#| fig-height: 4

p1 <- ggplot(penguins_clean, aes(x = species, y = flipper_length_mm, fill = species)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  labs(title = "Box Plot") +
  theme_minimal()

p2 <- ggplot(penguins_clean, aes(x = species, y = flipper_length_mm, fill = species)) +
  geom_violin(alpha = 0.5, show.legend = FALSE) +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1) +
  labs(title = "Violin + Jitter") +
  theme_minimal()

p1 + p2
```

### Line Plot

```{r}
#| label: line-plot
#| fig-width: 7
#| fig-height: 4

# Simulated time-course data
set.seed(123)
timecourse <- tibble(
  time  = rep(c(0, 6, 12, 24, 48), each = 3),
  gene  = rep(c("BRCA1", "TP53", "MYC"), times = 5),
  expr  = c(
    5, 12, 3,  7, 11, 5,  10, 9, 8,  12, 8, 12,  11, 7, 15
  )
)

ggplot(timecourse, aes(x = time, y = expr, color = gene)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title = "Gene Expression Time Course",
    x = "Time (hours)", y = "Expression (log2 CPM)"
  ) +
  theme_minimal()
```

### Heatmap — `geom_tile()`

```{r}
#| label: heatmap
#| fig-width: 6
#| fig-height: 5

# Correlation matrix of penguin numeric variables
cor_mat <- penguins_clean |>
  select(where(is.numeric)) |>
  cor() |>
  as.data.frame() |>
  rownames_to_column("var1") |>
  pivot_longer(-var1, names_to = "var2", values_to = "correlation")

ggplot(cor_mat, aes(x = var1, y = var2, fill = correlation)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(correlation, 2)), size = 3) +
  scale_fill_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0) +
  labs(title = "Correlation Heatmap") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 4.3 Aesthetics — Mapping vs Setting

```{r}
#| label: aesthetics
#| fig-width: 8
#| fig-height: 4

# MAPPING: variable controls the aesthetic
p1 <- ggplot(penguins_clean, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(color = species), size = 2) +
  labs(title = "Color mapped to species") +
  theme_minimal()

# SETTING: fixed value for all points
p2 <- ggplot(penguins_clean, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(color = "steelblue", size = 2, alpha = 0.6) +
  labs(title = "Color set to steelblue") +
  theme_minimal()

p1 + p2
```

### Custom Scales

```{r}
#| label: custom-scales
#| fig-width: 7
#| fig-height: 5

ggplot(penguins_clean, aes(x = body_mass_g, y = flipper_length_mm, color = species)) +
  geom_point(size = 2, alpha = 0.7) +
  scale_color_manual(values = c("Adelie" = "#E64B35", "Chinstrap" = "#4DBBD5", "Gentoo" = "#00A087")) +
  scale_x_continuous(labels = scales::comma) +
  theme_minimal()
```

## 4.4 Faceting

```{r}
#| label: facet-wrap
#| fig-width: 9
#| fig-height: 4

ggplot(penguins_clean, aes(x = bill_length_mm, y = body_mass_g)) +
  geom_point(aes(color = sex), alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.8) +
  facet_wrap(~species, scales = "free") +
  labs(title = "Bill Length vs Body Mass by Species") +
  theme_minimal()
```

```{r}
#| label: facet-grid
#| fig-width: 9
#| fig-height: 6

ggplot(penguins_clean, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(alpha = 0.5) +
  facet_grid(sex ~ species) +
  theme_bw()
```

## 4.5 Themes and Labels

```{r}
#| label: themes
#| fig-width: 7
#| fig-height: 5

ggplot(penguins_clean, aes(x = species, y = body_mass_g, fill = species)) +
  geom_boxplot(alpha = 0.8, show.legend = FALSE) +
  labs(
    title    = "Body Mass Distribution by Species",
    subtitle = "Palmer Penguins dataset",
    x        = NULL,
    y        = "Body Mass (g)",
    caption  = "Data: palmerpenguins R package"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_classic() +
  theme(
    plot.title    = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(color = "grey40"),
    axis.text     = element_text(size = 11)
  )
```

### Saving Plots

```{r}
#| label: ggsave
#| eval: false

# Save the last plot
ggsave("my_figure.png", width = 7, height = 5, dpi = 300)
ggsave("my_figure.pdf", width = 7, height = 5)
```


# Part 5: Statistical Analyses

```{r}
#| label: stats-setup

library(rstatix)
library(broom)
library(ggpubr)
```

## 5.1 Descriptive Statistics

```{r}
#| label: descriptive-stats

penguins_clean |>
  group_by(species) |>
  get_summary_stats(body_mass_g, type = "common")
```

### Normality Testing

```{r}
#| label: normality
#| fig-width: 7
#| fig-height: 4

# Shapiro-Wilk test per group
penguins_clean |>
  group_by(species) |>
  shapiro_test(body_mass_g)

# Visual check with QQ plot
ggplot(penguins_clean, aes(sample = body_mass_g)) +
  stat_qq() +
  stat_qq_line(color = "red") +
  facet_wrap(~species, scales = "free") +
  labs(title = "QQ Plots for Body Mass by Species") +
  theme_minimal()
```

## 5.2 T-Tests

### Two-Sample T-Test

```{r}
#| label: t-test

# Compare bill length between Adelie and Chinstrap
two_sp <- penguins_clean |>
  filter(species %in% c("Adelie", "Chinstrap"))

# Base R approach
t_result <- t.test(bill_length_mm ~ species, data = two_sp)
t_result

# Tidy output with broom
tidy(t_result)

# rstatix approach
two_sp |> t_test(bill_length_mm ~ species)
```

### Visualise with Significance

```{r}
#| label: t-test-plot
#| fig-width: 5
#| fig-height: 5

ggboxplot(two_sp, x = "species", y = "bill_length_mm",
          fill = "species", palette = "jco") +
  stat_compare_means(method = "t.test", label = "p.signif",
                     label.x = 1.5, label.y = 55)
```

## 5.3 ANOVA

### One-Way ANOVA

```{r}
#| label: anova

# Body mass across all three species
penguins_clean |> anova_test(body_mass_g ~ species)

# Post-hoc pairwise comparisons
penguins_clean |> tukey_hsd(body_mass_g ~ species)
```

### Check Assumptions

```{r}
#| label: anova-assumptions

# Levene's test for homogeneity of variances
penguins_clean |> levene_test(body_mass_g ~ species)
```

### Visualise ANOVA Results

```{r}
#| label: anova-plot
#| fig-width: 6
#| fig-height: 5

ggboxplot(penguins_clean, x = "species", y = "body_mass_g",
          fill = "species", palette = "npg") +
  stat_compare_means(method = "anova", label.y = 6800) +
  stat_compare_means(comparisons = list(
    c("Adelie", "Chinstrap"),
    c("Adelie", "Gentoo"),
    c("Chinstrap", "Gentoo")
  ), method = "t.test", label = "p.signif")
```

## 5.4 Correlation

```{r}
#| label: correlation

# Pearson correlation
penguins_clean |>
  cor_test(bill_length_mm, body_mass_g, method = "pearson")

# Correlation matrix
cor_results <- penguins_clean |>
  select(where(is.numeric)) |>
  cor_mat()
cor_results

# P-value matrix
cor_results |> cor_get_pval()
```

```{r}
#| label: correlation-plot
#| fig-width: 7
#| fig-height: 5

ggplot(penguins_clean, aes(x = bill_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species), alpha = 0.6) +
  geom_smooth(method = "lm", color = "black", linewidth = 0.8) +
  stat_cor(method = "pearson", label.x = 33, label.y = 6200) +
  labs(x = "Bill Length (mm)", y = "Body Mass (g)") +
  theme_minimal()
```

## 5.5 Linear Regression

### Simple Linear Regression

```{r}
#| label: simple-lm

model_simple <- lm(body_mass_g ~ bill_length_mm, data = penguins_clean)
summary(model_simple)
```

### Tidy Model Outputs with broom

```{r}
#| label: broom-outputs

# Coefficients
tidy(model_simple, conf.int = TRUE)

# Model-level statistics (R-squared, AIC, etc.)
glance(model_simple)

# Row-level predictions and residuals
augment(model_simple) |> head()
```

### Multiple Regression

```{r}
#| label: multiple-lm

model_multi <- lm(body_mass_g ~ bill_length_mm + flipper_length_mm + species,
                  data = penguins_clean)
tidy(model_multi, conf.int = TRUE)
glance(model_multi)
```

### Diagnostic Plots

```{r}
#| label: diagnostics
#| fig-width: 8
#| fig-height: 6

par(mfrow = c(2, 2))
plot(model_multi)
par(mfrow = c(1, 1))
```

## 5.6 Chi-Square Test

Test whether species distribution is independent of island:

```{r}
#| label: chi-square

# Contingency table
table(penguins_clean$species, penguins_clean$island)

# Chi-square test
chisq.test(table(penguins_clean$species, penguins_clean$island))

# Tidy approach
penguins_clean |>
  select(species, island) |>
  table() |>
  chisq_test()
```

```{r}
#| label: chi-square-plot
#| fig-width: 7
#| fig-height: 5

ggplot(penguins_clean, aes(x = island, fill = species)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Species Composition by Island",
    y = "Proportion"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()
```


# Summary

| Topic | Key Functions |
|-------|---------------|
| **Data types** | `class()`, `typeof()`, `is.*()`, `as.*()` |
| **Vectors** | `c()`, `seq()`, `rep()`, `length()`, `which()` |
| **Data frames** | `str()`, `summary()`, `head()`, `dim()` |
| **Read data** | `read_csv()`, `read_tsv()`, `read_excel()` |
| **Filter rows** | `filter()`, `slice_max()`, `distinct()` |
| **Select columns** | `select()`, `starts_with()`, `where()` |
| **Transform** | `mutate()`, `case_when()`, `across()` |
| **Summarize** | `summarize()`, `group_by()`, `count()` |
| **Reshape** | `pivot_longer()`, `pivot_wider()`, `separate()` |
| **Join** | `left_join()`, `inner_join()`, `anti_join()` |
| **Plot** | `ggplot()`, `geom_*()`, `facet_wrap()`, `theme_*()` |
| **T-test** | `t.test()`, `t_test()`, `stat_compare_means()` |
| **ANOVA** | `anova_test()`, `tukey_hsd()`, `levene_test()` |
| **Correlation** | `cor_test()`, `cor_mat()`, `stat_cor()` |
| **Regression** | `lm()`, `tidy()`, `glance()`, `augment()` |
| **Chi-square** | `chisq.test()`, `chisq_test()` |

## Next Steps

- **Bioconductor**: Explore packages for genomics (`DESeq2`, `edgeR`, `GenomicRanges`)
- **R for Data Science** (2e): Free online at [r4ds.hadley.nz](https://r4ds.hadley.nz)
- **Statistical Rethinking**: Bayesian approaches for scientific data analysis
- **Bioconductor Workflows**: Curated analysis pipelines at [bioconductor.org/packages/release/BiocViews.html#___Workflow](https://bioconductor.org/packages/release/BiocViews.html#___Workflow)
