---
title: "Stroke prevention: clustering and risks factors identification"
author: "William Guesdon"
output:
  html_document:
    code_folding: "hide"
    toc: true
    df_print: kable
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---

## Abstract
The goal of this project is to use statistical learning to identify the combination of the features that are more likely to be associated with stroke. For this analysis, I first performed an exploratory data analysis and feature engineering. As often seen in health-related datasets, the proportion of patients with stroke is low, resulting in a severely unbalanced dataset. To compensate for the unbalance, I used the ROSE package to artificially over sample the number of strokes observations.  
I then used a dimension reduction technique adapted to mixed datasets of continuous and categorical variables with the FAMD package. This visualization allowed me to identify the features most likely associated with the risk of developing a stroke.
Using these selected variables, I built a logistic regression model to evaluate the contribution of each feature to the risk of developing a stroke. The logistic regression model identified the Age, Average Glucose Level, Smoking, Hypertension, and Heart Diseases features as the most likely risk factors of developing a stroke. Age was the principal risk factor in this study.

_Any feedback is appreciated. If you like the notebook, please upvote._

## Dataset information and variables list  

* id: Patient ID
* gender: Gender of Patient
* age: Age of Patient
* hypertension: 0 - no hypertension, 1 hypertension
* heart-disease: 0 - no heart disease, 1 heart disease
* ever_married: Yes/No
* work_type: Type of occupation
* Residence_type: Are type of residence (Urban / Rural)
* avg_glucose_level: Average Glucose level (measure after meal)
* bmi: body mass index
* smoking_status: patient smoking status
* Stoke: 0 - no stroke, 1 - suffered stroke

## Initialization
```{r warning=FALSE, message=FALSE}
# Load libraries
library('kableExtra') # To print table in report
library('ggsci') # For color palettes 
library('glue') # For improved print
library('DataExplorer') # For Exploratory Data Analysis
library('patchwork') # To easily plot multiple graphs
library('ROSE') # For oversampling
library('FactoMineR') # For PCA Analysis
library('factoextra') # For PCA Analysis
library('tidyverse') # For Data Cleaning, EDA, Data Visualization
# Load dplyr last because other package can cause issue with the count function
```

```{r message=FALSE}
# Load the dataset
df <- read_csv("../input/healthcaredataset/train_2v.csv")
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Set upa  color blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

## Create output directory
```{r}
dir.create("../working/EDA")
dir.create("../working/Independent_variables")
dir.create("../working/Logistic_regression")
dir.create("../working/PCA")
dir.create("../working/Tables")
```

## Environement and packages informations
```{r}
# Print the version of the essential packages used for the analysis
packages_list <- c('tidyverse', 'DataExplorer', 'FactoMineR', 'factoextra', 'patchwork', 'caret', 'ROSE')

Name <- 'tidyverse'
Version <- getNamespaceVersion('tidyverse')

packages_informations <- tibble(Name, Version)

packages_list <- packages_list[2:length(packages_list)]
    
for(package in packages_list){
    Name <- package
    Version <- getNamespaceVersion(package)
    db <- tibble(Name, Version)
    packages_informations <- rbind(packages_informations, db)
}

packages_informations <- packages_informations %>% arrange(Name) 
write_csv(packages_informations, '../working/packages_informations.csv')
```

## Exploratory Data Analysis
### Check if the dataset is tidy
```{r}
df %>% head()
```

The dataset is tidy. Every column is a variable and every row is an observation.

### Is there missing value in the dataset?
```{r}
pdf('../working/EDA/plot_missing.pdf')        
plot_missing(df, ggtheme = theme_minimal())
dev.off()
```

Smoking is a known risk factor for stroke. Since smoking is a konow factor I will exclude patients with unknow status from this analysis as it could be confounding factor in the analysis. The number of missing value for this parameter is high wich is a concern for the analysis.
BMI is another know known risk factor so the ammount of missing value is alose a concern.

```{r}
# Strategy for missing values
df <- df %>% filter(!is.na(smoking_status)) # Exclude from the analysis patient with unknown smoking status
```

```{r}
df <- df %>% mutate(bmi = ifelse(is.na(bmi), median(bmi, na.rm=TRUE), bmi)) # replace missing bmi value by the median of the bmi
```

Only a few observation are labelled as others and I have no information on these patients so I will exclude them for the reamining of the analysis. 

```{r}
df <- df %>% filter(gender == 'Male' | gender == 'Female') # 
```

```{r}
df$id <- NULL # Remove the ID column

df <- df %>%
    mutate(
        stroke = ifelse(stroke == 0, 'no stroke', 'stroke'),
        hypertension = ifelse(hypertension == 0, 'no hypertension', 'hypertension'),
        heart_disease  = ifelse(heart_disease == 0, 'no heart disease', 'heart disease')
        )

# Trasnfrom categorical variable as factor
df$stroke <- as.factor(df$stroke)
df$gender <-  as.factor(df$gender)
df$ever_married <- as.factor(df$ever_married)
df$work_type <- as.factor(df$work_type)
df$Residence_type <- as.factor(df$Residence_type)
df$smoking_status <- as.factor(df$smoking_status)
df$hypertension <- as.factor(df$hypertension)
df$heart_disease <- as.factor(df$heart_disease)
```

### How many patients have stroke?
```{r}
total_observation <- nrow(df)
disease_tables <- df %>%
    group_by(stroke) %>%
    dplyr::count() %>% # Be carefull count() to not work because of MatrixStat package
    mutate(proportion = n/total_observation * 100) 
write_csv(disease_tables, '../working/Tables/disease_tables_df.csv')
```

As common with disease related data thre is a clear unbalance between disease and non disease patient wich is an issue for machine lerning methods.
I will use a strategy of oversampling to deal with his issue. 

### Are the gender balanced?
```{r}
# Calculate the table of gender repartition by disease status
total_observation <- nrow(df)
t1 <- df %>%
    group_by(gender, stroke) %>%
    dplyr::count() %>%
    mutate(
        percentage = n/total_observation * 100)
```

```{r}
# Summarize the age per gender and disease status
t2 <- df %>%
    group_by(gender, stroke) %>%
    summarise(
        age_mean = mean(age),
        age_sd = sd(age),
        n = n()
    )
```

```{r message=FALSE}
t3 <- inner_join(t1, t2)
t3
```

The proportion of disease per gender and age group is balanced.

### Visualize Data Distribution, Asses normality

```{r}
glue("{ncol(df)} variables and {nrow(df)} observatoins")
glue("Variable names:")
glue("{colnames(df)}")
```

```{r}
categorical_data <- c("gender", "hypertension", "heart_disease", "ever_married", "work_type", "Residence_type", "smoking_status", "stroke")
pdf('../working/EDA/categorical_variables.pdf')
plot_bar(df[categorical_data], ggtheme = theme_minimal())
dev.off()
```

```{r}
continous_data <- c("age", "avg_glucose_level", "bmi")
pdf('../working/EDA/continous_variables.pdf')
plot_density(df[continous_data], ggtheme = theme_minimal())
dev.off()
```

```{r}
pdf('../working/EDA/QQ_plots.pdf')
plot_qq(df[continous_data], ggtheme = theme_minimal(), sampled_rows = 100L)
dev.off()
```

```{r}
df <- df %>% mutate(bmi = log(bmi))

pdf('../working/EDA/Log_BMI_QQ_plot.pdf')
plot_qq(df[c('bmi')], ggtheme = theme_minimal(), sampled_rows = 100L)
dev.off()
```

The bmi distribution is right skewed wich can be solve by a log transformation.  
Average glucose level is bomodal and could potentialy be transformed as a categorical variable.  

## Solving Sampling Issue: ROSE

See [Lunardon, N., Menardi, G., and Torelli, N. (2014). ROSE: a Package for Binary Imbalanced Learning. R Journal, 6:82–92.](https://journal.r-project.org/archive/2014-1/menardi-lunardon-torelli.pdf)  

```{r}
# Over sample the data so stroke represent 50% of the total case.
data_oversampled <- ovun.sample(stroke ~ ., data = df, method = "over", N = 58940) # Increase the amount of stroke observation until it reach 29470, 58940 total observation
df_rose <- data_oversampled$data
# table(df_rose$stroke)
```

```{r}
total_observation <- nrow(df_rose)
disease_tables <- df_rose %>%
    group_by(stroke) %>%
    dplyr::count() %>% # Be carefull count() to not work because of MatrixStat package
    mutate(proportion = n/total_observation * 100) 
write_csv(disease_tables, '../working/Tables/disease_tables_df_rose.csv')
```

## FAMD - Factor Analysis of Mixed Data in R
See the [website](http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/115-famd-factor-analysis-of-mixed-data-in-r-essentials/) of the factoextra package creator: Dr. Alboukadel Kassambara. 

```{r}
res.famd <- FAMD(df_rose, graph = FALSE)
```

```{r message=FALSE}
p1 <- fviz_screeplot(res.famd)
ggsave('../working/PCA/scree_plot.png', p1)
```

```{r message=FALSE}
# Plot of variables
p1 <- fviz_famd_var(res.famd, repel = TRUE)
# Contribution to the first dimension
p2 <- fviz_contrib(res.famd, "var", axes = 1)
# Contribution to the second dimension
p3 <- fviz_contrib(res.famd, "var", axes = 2)

ggsave('../working/PCA/variables_FAMD.png', p1)

p4 <- (p2 / p3)
ggsave('../working/PCA/contribution_of_variables.png', p4)
plot(p4)
```

From this dimension reduction it appears that age, stroke and marrital status contribute the most to the dimension 1
smoking status, gender, work tupe and bmi contribute the mist to the dimesnion 2

```{r}
# quanti.var <- get_famd_var(res.famd, "quanti.var")
# quanti.var 
```

```{r message=FALSE}
p1 <- fviz_famd_var(res.famd, "quanti.var", col.var = "contrib", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE)
ggsave('../working/PCA/quantitative_variables.png', p1)
```

```{r}
quali.var <- get_famd_var(res.famd, "quali.var")
```

```{r message=FALSE}
p1 <- fviz_famd_var(res.famd, "quali.var", col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE
             )
ggsave('../working/PCA/qualitative_variables.png', p1)
plot(p1)
```

```{r message=FALSE}
famd.stroke <- fviz_mfa_ind(res.famd, 
             habillage = "stroke", # color by groups
             geom = c('point'),
             palette = pal_simpsons("springfield", alpha = 0.6)(16),
             title = "Stroke Status"
             ) 
ggsave('../working/PCA/famd.stroke.png', famd.stroke)
plot(famd.stroke)
```

```{r message=FALSE}
famd.smoking <- fviz_mfa_ind(res.famd, 
             habillage = "smoking_status", # color by groups
             geom = c('point'),
             palette = pal_simpsons("springfield", alpha = 0.6)(16),
             title = "Smoking Status"
             ) 
ggsave('../working/PCA/famd.smoking.png', famd.smoking)
```

```{r message=FALSE}
famd.married <- fviz_mfa_ind(res.famd, 
             habillage = "ever_married", # color by groups
             geom = c('point'),
             palette = pal_simpsons("springfield", alpha = 0.6)(16),
             title = "Martial Status"
             ) 
ggsave('../working/PCA/famd.married.png', famd.married)
```

```{r message=FALSE}
famd.work_type <- fviz_mfa_ind(res.famd, 
             habillage = "work_type", # color by groups
             geom = c('point'), 
             palette = pal_simpsons("springfield", alpha = 0.6)(16),
             title = "Work Status"
             ) 
ggsave('../working/PCA/famd.work_type.png', famd.work_type)
```

```{r message=FALSE}
famd.residence <- fviz_mfa_ind(res.famd, 
             habillage = "Residence_type", # color by groups
             geom = c('point'), # rempove labels
             palette = pal_simpsons("springfield", alpha = 0.6)(16), # use a color blind friendly palette 
             title = "Residence type"
             ) 
ggsave('../working/PCA/famd.residence.png', famd.residence)
```

```{r message=FALSE}
famd.heart_disease <- fviz_mfa_ind(res.famd, 
             habillage = "heart_disease", # color by groups
             geom = c('point'), 
             palette = pal_simpsons("springfield", alpha = 0.6)(16),
             title = "Heart Disease Status"
             )
ggsave('../working/PCA/famd.heart_disease.png', famd.heart_disease)
```

```{r message=FALSE}
famd.hypertension <- fviz_mfa_ind(res.famd, 
            habillage = "hypertension", # color by groups
            geom = c('point'),
            palette = pal_simpsons("springfield", alpha = 0.6)(16),
            title = "Hypertension Status"
             )
ggsave('../working/PCA/famd.hypertension.png', famd.hypertension)
```

```{r message=FALSE}
all_FAMD <- (famd.smoking + famd.married) / (famd.work_type + famd.residence) / (famd.heart_disease + famd.hypertension)
ggsave('../working/PCA/all_FAMD.png', all_FAMD)
plot(all_FAMD)
```

```{r message=FALSE}
famd.gender <- fviz_mfa_ind(res.famd, 
             habillage = "gender", # color by groups
             geom = c('point'),
             palette = pal_simpsons("springfield", alpha = 0.6)(16),
             title = "Gender"
             ) 
ggsave('../working/PCA/famd.gender.png', famd.gender)
```

## Stroke incidence vs age  
```{r message=FALSE}
p1 <- df_rose %>%
    ggplot(aes(y = stroke, x = age, fill = stroke)) +
    geom_boxplot(notch = TRUE) +     
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    guides(fill=FALSE) +
    ylab('') +
    ggtitle('Transformed dataset')

p2 <- df %>%
    ggplot(aes(y = stroke, x = age, fill = stroke)) +
    geom_boxplot(notch = TRUE) +     
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    guides(fill=FALSE) +
    ylab('') +
    ggtitle('Original dataset')

p3 <- p1 | p2
ggsave('../working/Independent_variables/stroke_vs_age.png', p3)
plot(p3)

age_summary <- df %>%
    group_by(stroke) %>%
    summarise(
        mean = mean(age),
        sd = sd(age),
        n = n()
    )

write_csv(age_summary, '../working/Independent_variables/age_summary.csv')
```

## Stroke incidence vs average glucose level       
```{r message=FALSE}
p1 <- df_rose %>%
    ggplot(aes(y = stroke, x = avg_glucose_level, fill = stroke)) +
    geom_boxplot(notch = TRUE) +     
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    guides(fill=FALSE) +
    ylab('') +
    xlab('Glucose Level') +
    ggtitle('Transformed dataset')


p2 <- df %>%
    ggplot(aes(y = stroke, x = avg_glucose_level, fill = stroke)) +
    geom_boxplot(notch = TRUE) +     
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    guides(fill=FALSE) +
    ylab('') +
    xlab('Glucose Level') +
    ggtitle('Original dataset')

p3 <- p1 | p2
ggsave('../working/Independent_variables/stroke_vs_avg_glucose_level.png', p3)
plot(p3)

avg_glucose_level_summary <- df %>%
    group_by(stroke) %>%
    summarise(
        mean = mean(avg_glucose_level),
        sd = sd(avg_glucose_level),
        n = n()
    )
write_csv(avg_glucose_level_summary, '../working/Independent_variables/avg_glucose_level_summary.csv')
```
     
## Stroke incidence vs gender
```{r message=FALSE}
p1 <- df_rose %>%
    ggplot(aes(x = gender, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Gender') +
    ggtitle('Transformed dataset')

p2 <- df %>%
    ggplot(aes(x = gender, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Gender') +
    ggtitle('Original dataset')

p3 <- p1 | p2
ggsave('../working/Independent_variables/stroke_vs_gender.png', p3)
plot(p3)
```

## Stroke incidence vs marital status
```{r message=FALSE}
p1 <- df_rose %>%
    ggplot(aes(x = ever_married, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Ever Married') +
    ggtitle('Transformed dataset')

p2 <- df %>%
    ggplot(aes(x = ever_married, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Ever Married') +
    ggtitle('Original dataset')

p3 <- p1 | p2
ggsave('../working/Independent_variables/stroke_vs_ever_married.png', p3)
plot(p3)
```

     
```{r message=FALSE}
p1 <- df_rose %>%
    ggplot(aes(y = ever_married, x = age, fill = ever_married)) +
    geom_boxplot(notch = TRUE) +     
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    guides(fill=FALSE) +
    ylab('Ever Married') +
    xlab('Age') +
    ggtitle('Transformed dataset')


p2 <- df %>%
    ggplot(aes(y = ever_married, x = age, fill = ever_married)) +
    geom_boxplot(notch = TRUE) +     
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    guides(fill=FALSE) +
    ylab('Ever Married') +
    xlab('Age') +
    ggtitle('Original dataset')

p3 <- p1 | p2
ggsave('../working/Independent_variables/stroke_vs_age.png', p3)
plot(p3)

age_summary <- df %>%
    group_by(ever_married) %>%
    summarise(
        mean = mean(age),
        sd = sd(age),
        n = n()
    )

write_csv(age_summary, '../working/Independent_variables/age_ever_married_summary.csv')
```

The impact of the marital status is likely a confunding factor due of the age difference.

 
 ## Stroke incidence vs smoking status    
```{r message=FALSE}
p1 <- df_rose %>%
    ggplot(aes(x = smoking_status, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Smoking Status') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Transformed dataset')

p2 <- df %>%
    ggplot(aes(x = smoking_status, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Smoking Status') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Original dataset')

p3 <- p1 | p2
ggsave('../working/Independent_variables/stroke_vs_smoking_status.png', p3)
plot(p3)
```
    
```{r message=FALSE}
p1 <- df_rose %>%
    ggplot(aes(y = smoking_status, x = age, fill = smoking_status)) +
    geom_boxplot(notch = TRUE) +     
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    guides(fill=FALSE) +
    ylab('Smoking Status') +
    xlab('Age') +
    ggtitle('Transformed dataset')


p2 <- df %>%
    ggplot(aes(y = smoking_status, x = age, fill = smoking_status)) +
    geom_boxplot(notch = TRUE) +     
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    guides(fill=FALSE) +
      ylab('Smoking Status') +
    xlab('Age') +
    ggtitle('Original dataset')

p3 <- p1 | p2
ggsave('../working/Independent_variables/stroke_vs_age.png', p3)
plot(p3)

age_summary <- df %>%
    group_by(ever_married) %>%
    summarise(
        mean = mean(age),
        sd = sd(age),
        n = n()
    )

write_csv(age_summary, '../working/Independent_variables/age_smoking_status_summary.csv')
```

Surprinsingly the subset who formerly snoked have an higher proportion of stroke. Having a look at the age distribution indicate that this could be caused by the fact that in this group subset who formerly skoked significantly older than the subject who smoke. To access the impact of every paramter in the study it is therefore important to randomize the age of the subsets.
     
```{r message=FALSE}
p1 <- df_rose %>%
    mutate(Ever_Smoked = ifelse((smoking_status == 'smokes' | smoking_status == 'formerly smoked'), 'Yes', 'No'))  %>%
    ggplot(aes(x = Ever_Smoked, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Ever Smoked') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Transformed dataset')

p2 <- df %>%
    mutate(Ever_Smoked = ifelse((smoking_status == 'smokes' | smoking_status == 'formerly smoked'), 'Yes', 'No'))  %>%
    ggplot(aes(x = Ever_Smoked, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Ever Smoked') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Original dataset')

p3 <- p1 | p2
plot(p3)
ggsave('../working/Independent_variables/stroke_vs_ever_smoked.png', p3)
```

 ## Stroke incidence vs hypertension
```{r message=FALSE}
p1 <- df_rose %>%
    ggplot(aes(x = hypertension, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Hypertension') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Transformed dataset')

p2 <- df %>%
    ggplot(aes(x = hypertension, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Hypertension') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Original dataset')

p3 <- p1 | p2
ggsave('../working/Independent_variables/stroke_vs_hypertension.png', p3)
```

```{r message=FALSE}
p1 <- df_rose %>%
    ggplot(aes(y = hypertension, x = age, fill = hypertension)) +
    geom_boxplot(notch = TRUE) +     
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    guides(fill=FALSE) +
    ylab('') +
    xlab('Age') +
    ggtitle('Transformed dataset')


p2 <- df %>%
    ggplot(aes(y = hypertension, x = age, fill = hypertension)) +
    geom_boxplot(notch = TRUE) +     
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    guides(fill=FALSE) +
    ylab('') +
    xlab('Age') +
    ggtitle('Original dataset')

p3 <- p1 | p2
ggsave('../working/Independent_variables/hypertension_vs_age.png', p3)
plot(p3)

write_csv(age_summary, '../working/Independent_variables/hypertension_summary.csv')
```

## Stroke incidence vs heart disease
```{r message=FALSE}
p1 <- df_rose %>%
    ggplot(aes(x = heart_disease, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Hypertension') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Transformed dataset')

p2 <- df %>%
    ggplot(aes(x = heart_disease, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Hypertension') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Original dataset')

p3 <- p1 | p2
ggsave('../working/Independent_variables/stroke_vs_heart_disease.png', p3)
plot(p3)
```

From the unsupervised learning is seems that the most likely associated features for a risk of developping a storke are: 

* Age
* Gender
* Average Glucose Level
* Ever Smoked
* Hypertension
* Heart Disease
     

## Find variable associated with stroke  - Logistic regression
There is many good explanation of logistic regression but my favorite are from [Josh Starmer](https://github.com/StatQuest) and DataCamp Instructor [Brett Lantz](https://www.amazon.co.uk/Machine-Learning-R-Brett-Lantz/dp/1782162143).

```{r}
df_selected <- df_rose %>%
    mutate(ever_smoked = ifelse((smoking_status == 'smokes' | smoking_status == 'formerly smoked'), 'Yes', 'No'))

df_selected <- df_selected %>% 
       select(age, gender, avg_glucose_level, ever_smoked, hypertension, heart_disease, stroke)

df_selected$ever_smoked <- as.factor(df_selected$ever_smoked)
```

```{r}
df_selected_features <- df_selected %>% select(-stroke) %>% mutate_if(is.factor, as.numeric) # Recode all categorical data to numeric
df_selected_target <- df_selected %>% select(stroke) 

df_selected <- bind_cols(df_selected_features, df_selected_target)

levels(df_selected$stroke)
```

```{r}
model <- glm(stroke ~ age + gender + avg_glucose_level + ever_smoked + hypertension + heart_disease, data = df_selected, family = binomial)
# exp(coef(model))
summary(model)
```

The deviances residual is centered on 0 and symetrical wich indicate that the model is likely accurate. 
The p values are significants for the most of the selected variable but gender. This not entirely surprising as the differences for gender were tiny on the bar plot. 

```{r}
## Now calculate the overall "Pseudo R-squared" and its p-value
ll.null <- model$null.deviance/-2
ll.proposed <- model$deviance/-2
 
## McFadden's Pseudo R^2 = [ LL(Null) - LL(Proposed) ] / LL(Null)
glue("McFadden's Pseudo R^2: {(ll.null - ll.proposed) / ll.null}")

## The p-value for the R^2
glue("R^2 p-value: {1 - pchisq(2*(ll.proposed - ll.null), df=(length(model$coefficients)-1))}")

```

A McFadden's Pseudo R^2 between 0.2 and 0.4 is considered a good fit.
Since the p-value is inferior to p=0.05 we reject the null hypothesis and consider the R^2 score significant.

```{r}
predicted.data <- data.frame(
  probability.of.stroke=model$fitted.values,
  stroke=df_selected$stroke)
```

```{r}
predicted.data <- predicted.data[
  order(predicted.data$probability.of.stroke, decreasing=FALSE),]
predicted.data$rank <- 1:nrow(predicted.data)
```

```{r message=FALSE}
p1 <- ggplot(data=predicted.data, aes(x=rank, y=probability.of.stroke)) +
  geom_point(aes(color=stroke), alpha=1, shape=4, stroke=2) +
  scale_colour_manual(values=cbPalette) +
  xlab("Index") +
  ylab("Predicted probability of getting a stroke") +
  theme_bw()
ggsave('../working/Logistic_regression/logisitc_regression_evaluation.png', p1)
plot(p1)
```

This graph ilustrate that most of the the stroke subjects are correctly callsified validation the model and therefore the identification of the features associated to stroke risk.

## Visualize the risk of stroke for the 5 risk factors identified  

```{r message=FALSE, fig.height=15, fig.width=5}
# Save a plot of the 5 selected risk factors from the transformed data frame

p1 <- df_rose %>%
    ggplot(aes(y = stroke, x = age, fill = stroke)) +
    geom_boxplot(notch = TRUE) +     
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    guides(fill=FALSE) +
    xlab('Age') +
    ylab('Stroke') +
    ggtitle('Age')

p2 <- df_rose %>%
    ggplot(aes(y = stroke, x = avg_glucose_level, fill = stroke)) +
    geom_boxplot(notch = TRUE) +     
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    guides(fill=FALSE) +
    ylab('Stroke') +
    xlab('Avegrage glucose Level') +
    ggtitle('Glucose Level')

p3 <- df_rose %>%
    mutate(Ever_Smoked = ifelse((smoking_status == 'smokes' | smoking_status == 'formerly smoked'), 'Yes', 'No'))  %>%
    ggplot(aes(x = Ever_Smoked, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Ever Smoked') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Smoking history')

p4 <-  df_rose %>%
    ggplot(aes(x = hypertension, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Hypertension') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Hypertension')

p5 <- df_rose %>%
    ggplot(aes(x = heart_disease, fill = stroke)) +
    geom_bar(position = 'fill') +
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    xlab('Heart Disesae') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Heart Disesae History')

p6 <- df_rose %>%
    ggplot(aes(y = stroke, x = bmi, fill = stroke)) +
    geom_boxplot(notch = TRUE) +     
    scale_fill_manual(values=cbPalette) +
    theme_bw() +
    guides(fill=FALSE) +
    xlab('BMI') +
    ylab('Stroke') +
    ggtitle('BMI')

p7 <- (p1 + p2) / (p3 + p4) / (p5 + plot_spacer())
ggsave('../working/Independent_variables/risks_factors.png', p7)
plot(p7)
```

## Conclusion
Using dimension reduction and supervised machine learning, I was able to identify 5 variables as the principal risk factors of developing a stroke: 
  
* Age
* Average Glucose Level
* Ever Smoked
* Hypertension
* Heart Disease

This features could be used to built a machine learning model of stroke risk using the logistic regression or random forest. Although given the limited number of initial observations such model is more at risk to over fit even after correction for sample unbalanced.

## References
* [Healthcare Dataset Stroke Data](https://www.kaggle.com/asaumya/healthcare-dataset-stroke-data)
* [Lunardon, N., Menardi, G., and Torelli, N. (2014). ROSE: a Package for Binary Imbalanced Learning. R Journal, 6:82–92.](https://journal.r-project.org/archive/2014-1/menardi-lunardon-torelli.pdf)
* [FAMD - Factor Analysis of Mixed Data in R: Essentials](http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/115-famd-factor-analysis-of-mixed-data-in-r-essentials/)
* [logistic_regression_demo](https://github.com/StatQuest/logistic_regression_demo/blob/master/logistic_regression_demo.R)
* [Machine Learning with R - Brett Lantz](https://www.amazon.co.uk/Machine-Learning-techniques-predictive-modeling-ebook/dp/B07PYXX3H5/ref=sr_1_1?dchild=1&qid=1585504635&refinements=p_27%3ABrett+Lantz&s=books&sr=1-1&text=Brett+Lantz)