---
title: "Heart Disease: EDA and Random forest with R"  
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'  
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE,
	paged.print=TRUE
)
```
## Introduction
It is my first kernel on the community so any feedback would be much appreciated.  
If my work is useful to you, please upvote the code.  
Thanks!  

## Load libraries  
* tidyverse: For data cleaning, sorting, and visualization  
* DataExplorer: For Exploratory Data Analysis  
* gridExtra: To plot several plots in one figure  
* ggpubr: To prepare publication-ready plots  
* GGally: For correlations  
* caTools: For classification model 
* rpart: For classification model  
* rattle: Plot nicer descision trees  
* randomForest: For Random Forest model  

```{r message = FALSE, warning = FALSE}
library(tidyverse)  # For data cleaning, sorting, and visualization
library(DataExplorer) # For Exploratory Data Analysis
library(gridExtra) # To plot several plots in one figure
library(ggpubr) # To prepare publication-ready plots
library(GGally) # For correlations
library(caTools) # For classification model
library(rpart) # For classification model
library(rattle) # Plot nicer descision trees
library(randomForest) # For Random Forest model
```

## I Data


heart-disease-uci Kaggle 
https://www.kaggle.com/ronitf/heart-disease-uci   

Useful links for this dataset  
https://lucdemortier.github.io/projects/3_mcnulty   
https://www.kaggle.com/ronitf/heart-disease-uci/discussion/105877  
https://www.kaggle.com/ronitf/heart-disease-uci/discussion/93372  

**As explained on the links above, it is essential to note that on this dataset, the target value 0 indicates that the patient has heart disease.**

Attribute Information:  
age: age in years  
sex: (1 = male; 0 = female)  
cp: chest pain type  (typical angina, atypical angina, non-angina, or asymptomatic angina)  
trestbps: resting blood pressure (in mm Hg on admission to the hospital)  
chol: serum cholestoral in mg/dl  
fbs: Fasting blood sugar (< 120 mg/dl or > 120 mg/dl) (1 = true; 0 = false)  
restecg: resting electrocardiographic results (normal, ST-T wave abnormality, or left ventricular hypertrophy)  
thalach: Max. heart rate achieved during thalium stress test  
exang: Exercise induced angina (1 = yes; 0 = no)  
oldpeak: ST depression induced by exercise relative to rest  
slope: Slope of peak exercise ST segment (0 = upsloping, 1 = flat, or 2 = downsloping)  
ca: number of major vessels (0-3) colored by flourosopy  4 = NA  
thal: Thalium stress test result 3 = normal; 6 = fixed defect; 7 = reversable defect  0 = NA  
target: Heart disease status 1 or 0 (0 = heart disease 1 = asymptomatic) 

```{r message = FALSE}
df <- read_csv("./Data/heart.csv") # To read file on Kaggle
```

## II Tidy dataset

```{r}
copy <- df

df2 <- df %>%
  filter(
    thal != 0 & ca != 4 # remove values correspondind to NA in original dataset
  ) %>%
# Recode the categorical variables as factors using the dplyr library.
  mutate(
    sex = case_when(
      sex == 0 ~ "female",
      sex == 1 ~ "male"
           ),
    fbs = case_when(
      fbs == 0 ~ "<=120",
      fbs == 1 ~ ">120"
            ),
    exang = case_when(
      exang == 0 ~ "no",
      exang == 1 ~ "yes"
            ),
    cp = case_when(
      cp == 3 ~ "typical angina",
      cp == 1 ~ "atypical angina",
      cp == 2 ~ "non-anginal",
      cp == 0 ~ "asymptomatic angina"
          ),
    restecg = case_when(
      restecg == 0 ~ "hypertrophy",
      restecg == 1 ~ "normal",
      restecg == 2 ~ "wave abnormality"
              ),
    target = case_when(
      target == 1 ~ "asymptomatic",
      target == 0 ~ "heart-disease"
              ),
    slope = case_when(
      slope == 2 ~ "upsloping",
      slope == 1 ~ "flat",
      slope == 0 ~ "downsloping"
    ),
    thal = case_when(
      thal == 1 ~ "fixed defect",
      thal == 2 ~ "normal",
      thal == 3 ~ "reversable defect"
    ),
    sex = as.factor(sex),
    fbs = as.factor(fbs),
    exang = as.factor(exang),
    cp = as.factor(cp),
    slope = as.factor(slope),
    ca = as.factor(ca),
    thal = as.factor(thal)
  )

glimpse(df2) # Check that the transformnation worked
plot_missing(df2) # Check that the transformation did not induce NA values
df <- df2 # Replace the df dataset by the tidy dataset
```

## III Exploratory Data Analysis
### A Visualize the data summary and distribution of each variable

```{r}
df %>%
  summary()
```

Use the DataExplorer library to get a sense of the distribution of the continuous and categorical variables.

```{r}
plot_density(df, ggtheme = theme_classic2(), geom_density_args = list("fill" = "black", "alpha" = 0.6))
plot_bar(df, ggtheme = theme_classic2())
```

The next step is to combine dplyr and Data Explorer libraries to visualize the variables according to gender and disease.  

### B Analyze each variable per gender

```{r}
df %>%
  filter(sex == "female") %>%
  plot_density(ggtheme = theme_classic2(), geom_density_args = list("fill" = "black", "alpha" = 0.6))
df %>%
  filter(sex == "male") %>%
  plot_density(ggtheme = theme_classic2(), geom_density_args = list("fill" = "black", "alpha" = 0.6))
```

```{r}
df %>%
  filter(sex == "female") %>%
  plot_bar(ggtheme = theme_classic2())
df %>%
  filter(sex == "male") %>%
  plot_bar(ggtheme = theme_classic2())
```

### C Visualize variables per disease status

```{r}
df %>%
  filter(target == "asymptomatic") %>%
  plot_density(ggtheme = theme_classic2(), geom_density_args = list("fill" = "black", "alpha" = 0.6))
df %>%
  filter(target == "heart-disease") %>%
  plot_density(ggtheme = theme_classic2(), geom_density_args = list("fill" = "black", "alpha" = 0.6))
```

```{r}
df %>%
  filter(target == "asymptomatic") %>%
  plot_bar(ggtheme = theme_classic2())
df %>%
  filter(target == "heart-disease") %>%
  plot_bar(ggtheme = theme_classic2())
```

### Visualize the data per gender and disease status
```{r}
df %>%
  filter(sex == "female", target == "asymptomatic") %>%
  plot_density(ggtheme = theme_classic2(), geom_density_args = list("fill" = "black", "alpha" = 0.6))
df %>%
  filter(sex == "female", target == "heart-disease") %>%
  plot_density(ggtheme = theme_classic2(), geom_density_args = list("fill" = "black", "alpha" = 0.6))
```

```{r}
df %>%
  filter(sex == "female", target == "asymptomatic") %>%
  plot_bar(ggtheme = theme_classic2())
df %>%
  filter(sex == "female", target == "heart-disease") %>%
  plot_bar(ggtheme = theme_classic2())
```

```{r}
df %>%
  filter(sex == "male", target == "asymptomatic") %>%
  plot_density(ggtheme = theme_classic2(), geom_density_args = list("fill" = "black", "alpha" = 0.6))
df %>%
  filter(sex == "male", target == "heart-disease") %>%
  plot_density(ggtheme = theme_classic2(), geom_density_args = list("fill" = "black", "alpha" = 0.6))
```

```{r}
df %>%
  filter(sex == "male", target == "asymptomatic") %>%
  plot_bar(ggtheme = theme_classic2())
df %>%
  filter(sex == "male", target == "heart-disease") %>%
  plot_bar(ggtheme = theme_classic2())
```

### Prepare a summary table per disease and gender
```{r}
df %>%
  group_by(target, sex) %>%
  summarise(
    n_disease = n(),
    mean_age = round(mean(age), digits=2),
    sd_age = round(sd(age), digits=2),
    mean_trestbps = round(mean(trestbps), digits=2),
    sd_trestbps = round(sd(trestbps), digits=2),
    mean_chol = round(mean(chol), digits=2),
    sd_chol = round(sd(chol), digits=2),
    mean_thalach = round(mean(thalach), digits=2),
    sd_thalach = round(sd(thalach), digits=2),
    mean_oldpeak = round(mean(oldpeak), digits=2),
    sd_oldpeak = round(sd(oldpeak), digits=2)
  )
```

## IV Data Visualization

From the Exploratory Data analysis, it seems that several differences are statistically significant according to gender and health status. 

### A Visualization of variables per gender

```{r}
# Male and Female count
a1 <- ggplot(df, aes(x = sex, fill = sex)) +
  geom_bar(width = 0.5) + 
  scale_fill_manual(values = c("#386cb0","#fdb462"))+
  theme_classic2() +
  theme(legend.position='none')

# Age per gender
b1 <- ggplot(df, aes(x= sex, y = age, fill = sex)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) +
  ylim(0, 90) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#386cb0","#fdb462"))+
  theme_classic2() +
  theme(legend.position='none')

# trestbps
c1 <- ggplot(df, aes(x = sex, y = trestbps, fill = sex)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) + 
  labs(y = "blood pressure (mmHg)") +
  ylim(0,250) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#386cb0","#fdb462"))+
  theme_classic2() +
  theme(legend.position='none')

# chol
d1 <- ggplot(df, aes(x = sex, y = chol, fill = sex)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) + 
  labs(y = "cholestorol (mg/dl)") +
  ylim(0,500) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#386cb0","#fdb462"))+
  theme_classic2() +
  theme(legend.position='none')

# oldpeak
e1 <- ggplot(df, aes(x = sex, y = oldpeak, fill = sex)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) + 
  labs(y = "ST depression") +
  ylim(0,10) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#386cb0","#fdb462"))+
  theme_classic2() +
  theme(legend.position='none')

# thalach
f1 <- ggplot(df, aes(x = sex, y = thalach, fill = sex)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) + 
  labs(y = "Max. heart rate") +
  ylim(0,250) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#386cb0","#fdb462"))+
  theme_classic2() +
  theme(legend.position='none')

suppressWarnings(ggarrange(a1, b1, c1, d1, e1, f1, 
          ncol = 2, nrow = 3,
            align = "v"))
```

```{r}
# Disease status
g1 <- ggplot(df, aes(x = target, fill = sex)) +
  geom_bar(width = 0.5, position = 'dodge') + 
  labs(x = "") +
  coord_flip() +
  scale_fill_manual(values = c("#386cb0","#fdb462"))+
  theme_classic2() +
  theme(legend.position='none')

# cp
h1 <- ggplot(df, aes(cp, group = sex, fill = sex)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "chest pain") +
  coord_flip() +
  scale_fill_manual(values = c("#386cb0","#fdb462"))+
  theme_classic2() +
  theme(legend.position='none')

# restecg
i1 <- ggplot(df, aes(restecg, group = sex, fill = sex)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "rest. electrocardiographic") +
  coord_flip() +
  scale_fill_manual(values = c("#386cb0","#fdb462"))+
  theme_classic2() +
  theme(legend.position='none')

# slope
j1 <- ggplot(df, aes(slope, group = sex, fill = sex)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "peak exercise ST") +
  coord_flip() +
  scale_fill_manual(values = c("#386cb0","#fdb462"))+
  theme_classic2() +
  theme(legend.position='none')

# thal 
k1 <- ggplot(df, aes(thal, group = sex, fill = sex)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "Thalium stress test") +
  coord_flip() +
  scale_fill_manual(values = c("#386cb0","#fdb462"))+
  theme_classic2() +
  theme(legend.position='none')

# fbp
l1 <- ggplot(df, aes(fbs, group = sex, fill = sex)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "Fasting blood sugar") +
  coord_flip() +
  scale_fill_manual(values = c("#386cb0","#fdb462"))+
  theme_classic2() +
  theme(legend.position='none')

# exang
m1 <- ggplot(df, aes(exang, group = sex, fill = sex)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "Exercise induced angina") +
  coord_flip() +
  scale_fill_manual(values = c("#386cb0","#fdb462"))+
  theme_classic2() +
  theme(legend.position='none')

# ca
n1 <- ggplot(df, aes(ca, group = sex, fill = sex)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "flourosopy") +
  coord_flip() +
  scale_fill_manual(values = c("#386cb0","#fdb462"))+
  theme_classic2() +
  theme(legend.position='none')

ggarrange(g1, h1, i1, j1, k1, l1, m1, n1, 
          ncol = 2, nrow = 4,
          align = "v")
```

**From this first plot, it appears that this dataset contains more males patients with a higher proportion of heart disease compared to female patients.**

### B Visualization of variables per disease status
#### 1 Male patient

```{r}
df <- df2 %>%
    filter(sex == "male")
```

```{r}
# Male and Female count
a2 <- ggplot(df, aes(x = target, fill = target)) +
  geom_bar(width = 0.5, position = 'dodge') + 
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# Age per gender
b2 <- ggplot(df, aes(x= target, y = age, fill = target)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) +
  ylim(0, 90) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# trestbps
c2 <- ggplot(df, aes(x = target, y = trestbps, fill = target)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) + 
  labs(y = "blood pressure (mmHg)") +
  ylim(0,250) +
 stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# chol
d2 <- ggplot(df, aes(x = target, y = chol, fill = target)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) + 
  labs(y = "cholestorol (mg/dl)") +
  ylim(0,500) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# oldpeak
e2 <- ggplot(df, aes(x = target, y = oldpeak, fill = target)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) + 
  labs(y = "ST depression") +
  ylim(0,10) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# thalach
f2 <- ggplot(df, aes(x = target, y = thalach, fill = target)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) + 
  labs(y = "Max. heart rate") +
  ylim(0,250) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

ggarrange(a2, b2, c2, d2, e2, f2, 
          ncol = 2, nrow = 3,
            align = "v")
```

**Male patients with heart disease are significantly older, have higher cholesterol level, and reduced maximum heart rate response to the thallium test. **

```{r}
# Disease status
g2 <- ggplot(df, aes(x = target, fill = target)) +
  geom_bar(width = 0.5, position = 'dodge') + 
  labs(x = "") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# cp
h2 <- ggplot(df, aes(cp, group = target, fill = target)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "chest pain") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# restecg
i2 <- ggplot(df, aes(restecg, group = target, fill = target)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "rest. electrocardiographic") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# slope
j2 <- ggplot(df, aes(slope, group = target, fill = target)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "peak exercise ST") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# thal 
k2 <- ggplot(df, aes(thal, group = target, fill = target)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "Thalium stress test") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# fbp
l2 <- ggplot(df, aes(fbs, group = target, fill = target)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "Fasting blood sugar") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# exang
m2 <- ggplot(df, aes(exang, group = target, fill = target)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "Exercise induced angina") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# ca
n2 <- ggplot(df, aes(ca, group = target, fill = target)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "flourosopy") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

ggarrange(g2, h2, i2, j2, k2, l2, m2, n2, 
          ncol = 2, nrow = 4,
          align = "v")
```

#### 2 Female patients

```{r}
df <- df2 %>%
    filter(sex == "female")
```

```{r}
# Male and Female count
a2 <- ggplot(df, aes(x = target, fill = target)) +
  geom_bar(width = 0.5, position = 'dodge') + 
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# Age per gender
b2 <- ggplot(df, aes(x= target, y = age, fill = target)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) +
  ylim(0, 90) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# trestbps
c2 <- ggplot(df, aes(x = target, y = trestbps, fill = target)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) + 
  labs(y = "blood pressure (mmHg)") +
  ylim(0,250) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# chol
d2 <- ggplot(df, aes(x = target, y = chol, fill = target)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) + 
  labs(y = "cholestorol (mg/dl)") +
  ylim(0,500) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# oldpeak
e2 <- ggplot(df, aes(x = target, y = oldpeak, fill = target)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) + 
  labs(y = "ST depression") +
  ylim(0,10) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# thalach
f2 <- ggplot(df, aes(x = target, y = thalach, fill = target)) +
  geom_violin(width = 0.5) +
  geom_boxplot(width = 0.2) + 
  labs(y = "Max. heart rate") +
  ylim(0,250) +
  stat_compare_means(aes(label = ..p.signif..), method = "t.test") +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

suppressWarnings(ggarrange(a2, b2, c2, d2, e2, f2, 
          ncol = 2, nrow = 3,
            align = "v"))
```

**There is less woman with heart disease on this data set. 
Women with heart disease have a significantly higher resting blood presure contrary to male with heart disease.
Similarly to men, women with heart disease have a lower maximum heart rate in response to the thallium test. 
**

```{r}
# Disease status
g2 <- ggplot(df, aes(x = target, fill = target)) +
  geom_bar(width = 0.5, position = 'dodge') + 
  labs(x = "") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# cp
h2 <- ggplot(df, aes(cp, group = target, fill = target)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "chest pain") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# restecg
i2 <- ggplot(df, aes(restecg, group = target, fill = target)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "rest. electrocardiographic") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# slope
j2 <- ggplot(df, aes(slope, group = target, fill = target)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "peak exercise ST") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# thal 
k2 <- ggplot(df, aes(thal, group = target, fill = target)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "Thalium stress test") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# fbp
l2 <- ggplot(df, aes(fbs, group = target, fill = target)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "Fasting blood sugar") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# exang
m2 <- ggplot(df, aes(exang, group = target, fill = target)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "Exercise induced angina") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

# ca
n2 <- ggplot(df, aes(ca, group = target, fill = target)) +
  geom_bar(position = "dodge") +
  labs(x = "", y = "flourosopy") +
  coord_flip() +
  scale_fill_manual(values = c("#7fc97f","#ef3b2c"))+
  theme_classic2() +
  theme(legend.position='none')

ggarrange(g2, h2, i2, j2, k2, l2, m2, n2, 
          ncol = 2, nrow = 4,
          align = "v")
```

## V Correlations
### A use the numerical dataset
```{r}
df <- copy %>%
  filter(
    thal != 0 & ca != 4 # remove values correspondind to NA in original dataset
    )
  
```

```{r}
# ggcorr(df, palette = "RdBu")
GGally::ggcorr(df, geom = "circle")
```

```{r}
select2 <- df %>%
  dplyr::select(
    target,
    slope,
    thalach,
    restecg,
    cp
  )
```

```{r}
ggcorr(select2, geom = "circle")
```

```{r}
ggpairs(df)
```

```{r}
ggpairs(select2)
```

From the correlation study it seems that the parameters    
* cp  
* restecg  
* thalach  
* slope   
are the most usefull to predict the risk of heart disease  

From the EDA anlysis it semms that  
* age  
* sex  
* cholesterol  
* restecg   
are also usefull  

For prediction the following variables seems the most usefull  
* age  
* sex  
* cholesterol  
* restecg  
* cp  
* thalach  
* slope  

## VI Machine Learning: classification model with rpart and random forest packages
1. Select the columns usefull for prediction according to the EDA analysis.  
2. Separate the data set in a train and test subsets.  
3. Build a classification tree model with rpart.  
4. Print model accuracy and descision tree.  

### A Use select columns for classification
```{r}
df_select <- df %>%
  dplyr::select( #because of conflict between MASS and dplyr select need to use dplyr::select
    target,
    age,
    sex,
    chol,
    restecg,
    cp,
    thalach,
    slope
  )
```

```{r}
df_select$target <- factor(df_select$target) # Define target as a factor. rpart classification would not work otherwise.

accuracy <- 0

# Build a simple classification desicion tree with rpart. Run the model until the accuracy reach the selected minimum.
while(accuracy <= 0.85) {
  split_values <- sample.split(df_select$target, SplitRatio = 0.65)
  train_set <- subset(df_select, split_values == T)
  test_set <- subset(df_select, split_values == F)
  mod_class <- rpart(target~. , data=train_set)
  result_class <- predict(mod_class, test_set, type = "class")
  table <- table(test_set$target, result_class)
  accuracy <- (table["0","0"] + table["1","1"])/sum(table)
  # cat("accuracy = ", round(accuracy, digits = 2)*100, "%")
}
```

Print model accuracy.  
According to parameters the model should be at least 88% accurate.  
```{r}
cat("Model accuracy", round(accuracy, digits = 2)*100, "%")  
```

Print the desicion tree.  
```{r}
# par(mfrow = c(1,2), xpd = NA) # otherwise on some devices the text is clipped
fancyRpartPlot(mod_class, , caption = NULL)
# plot(mod_class)
# text(mod_class, use.n = TRUE)
```

### B Use the full dataset for classification
```{r}
copy2 <- df
df$target <- factor(df$target)
accuracy <- 0

# Build a simple classification desicion tree with rpart. Run the model until the accuracy reach the selected minimum.
while(accuracy <= 0.88) {
  split_values <- sample.split(df_select$target, SplitRatio = 0.65)
  train_set <- subset(df, split_values == T)
  test_set <- subset(df, split_values == F)
  mod_class <- rpart(target~. , data=train_set)
  result_class <- predict(mod_class, test_set, type = "class")
  table <- table(test_set$target, result_class)
  accuracy <- (table["0","0"] + table["1","1"])/sum(table)
  # cat("accuracy = ", round(accuracy, digits = 2)*100, "%")
}
```

Print model accuracy.  
According to parameters the model should be at least 88% accurate.  
```{r}
cat("Model accuracy", round(accuracy, digits = 2)*100, "%")  
```

Print the desicion tree.  
```{r}
# par(mfrow = c(1,2), xpd = NA) # otherwise on some devices the text is clipped
fancyRpartPlot(mod_class, , caption = NULL)
# plot(mod_class)
# text(mod_class, use.n = TRUE)
```

### C Prediction on selected column with random forest
```{r}
set.seed(123)
train <- sample(nrow(df_select), 0.7*nrow(df_select), replace = FALSE)
TrainSet <- df_select[train,]
ValidSet <- df_select[-train,]
summary(TrainSet)
summary(ValidSet)
```

```{r}
# Create a Random Forest model with default parameters
model1 <- randomForest(target ~ ., data = TrainSet,  ntree = 1000, mtry = 1, importance = TRUE)
model1
```

```{r}
# Predicting on train set
predTrain <- predict(model1, TrainSet, type = "class")
# Checking classification accuracy
table(predTrain, TrainSet$target) 
```

```{r}
# Predicting on Validation set
predValid <- predict(model1, ValidSet, type = "class")
# Checking classification accuracy
mean(predValid == ValidSet$target)                    
table(predValid,ValidSet$target)
```

```{r}
# To check important variables
importance(model1)        
varImpPlot(model1)  
```


### D Use the full dataset for classification with random forest
```{r}
set.seed(123)
train <- sample(nrow(df), 0.7*nrow(df_select), replace = FALSE)
TrainSet <- df[train,]
ValidSet <- df[-train,]
summary(TrainSet)
summary(ValidSet)
```

```{r}
# Create a Random Forest model with default parameters
model2 <- randomForest(target ~ ., data = TrainSet,  ntree = 1000, mtry = 2, importance = TRUE)
model2
```

```{r}
# Predicting on train set
predTrain <- predict(model2, TrainSet, type = "class")
# Checking classification accuracy
table(predTrain, TrainSet$target) 
```

```{r}
# Predicting on Validation set
predValid <- predict(model2, ValidSet, type = "class")
# Checking classification accuracy
mean(predValid == ValidSet$target)                    
table(predValid,ValidSet$target)
```

```{r}
# To check important variables
importance(model2)        
varImpPlot(model2)  
```

## References
### Data transformation  
* https://archive.ics.uci.edu/ml/datasets/Heart+Disease  
* https://lucdemortier.github.io/projects/3_mcnulty  
* https://www.kaggle.com/ronitf/heart-disease-uci/discussion/105877  

### Kaggles notebooks:  
* R notebooks:  
* https://www.kaggle.com/ekrembayar/heart-disease-uci-eda-models-with-r  
* https://www.kaggle.com/joemenifee/heart-disease-uci-data-exploratory  

### Data Processing  
* http://www.cookbook-r.com/
* https://bookdown.org/rdpeng/exdata/managing-data-frames-with-the-dplyr-package.html#data-frames 
* https://rpkgs.datanovia.com/ggpubr/reference/stat_compare_means.html  
* https://towardsdatascience.com/simple-fast-exploratory-data-analysis-in-r-with-dataexplorer-package-e055348d9619  
* http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/81-ggplot2-easy-way-to-mix-multiple-graphs-on-the-same-page/  
* https://www.r-graph-gallery.com/267-reorder-a-variable-in-ggplot2  
* for categorical variable  
* https://www.analyticsvidhya.com/blog/2015/11/easy-methods-deal-categorical-variables-predictive-modeling/  
for correlations  
* http://www.sthda.com/english/wiki/ggally-r-package-extension-to-ggplot2-for-correlation-matrix-and-survival-plots-r-software-and-data-visualization  

### for Machine Learning  
* https://www.kaggle.com/naik170106027/prediction-of-heart-diseases  
* https://www.kaggle.com/ekrembayar/heart-disease-uci-eda-models-with-r  
* https://www.kaggle.com/anirbanshaw24/heart-disease-prediction-and-indicators  
* http://topepo.github.io/caret/index.html  
* https://www.youtube.com/watch?v=SeyghJ5cdm4&feature=youtu.be  
* https://www.rdocumentation.org/packages/rpart/versions/4.1-15/topics/rpart  
* https://stackoverflow.com/questions/33767804/invalid-prediction-for-rpart-object-error  
* https://machinelearningmastery.com/overfitting-and-underfitting-with-machine-learning-algorithms/  
* https://www.gormanalysis.com/blog/decision-trees-in-r-using-rpart/  
* https://www.kaggle.com/wguesdon/tuning-random-forest-parameters/edit  

### For Rstudio  
* https://community.rstudio.com/t/shortcut-to-remove-results-from-r-notebook/9147  